|

### 2019-10-23 manuscript/manuscript.Rmd
additions

---
title: |
 Towards Explainability of Machine Learning Models in Insurance Pricing
type: WIP DRAFT 
author:
  - name: Author One
    affil: a
    email: foo@bar.foo
  - name: Author Two
    affil: b
    email: bar@foo.bar
affiliation:
  - num: a
    address: |
      Foo
  - num: b
    address: |
      Bar
bibliography: explain-pricing.bib
appendix: appendix.tex
abstract: |
  Abstract tbd
keywords: |
  actuarial science; general insurance
header-includes: |
  \usepackage{hyperref}
  \usepackage[utf8]{inputenc}
  \def\tightlist{}
output: rticles::tf_article
---

@@

## Not All Linear Models are Interpretable

@@

Proposed by @hiltonConversationalProcesses1990 and interpreted by @millerExplanationArtificial2017 in the context of ML, model explanations are conversations or social interactions. One consequence of this identification is that explanations need to be *relevant* to the *audience*. In developing, filing, and operationalizing a pricing model, one needs to accommodate a variety of stakeholders, each of whom has a different set of questions, assumptions, and technical capacity. First, there are internal stakeholders at the company, which includes management and underwriters. While some of the individuals in this audience may be technical, they are likely less familiar with predictive modeling techniques than the actuaries and data scientists who build the models. Next, we have the regulators, who may have limited resources to review the models, and will focus on a specific list of questions. Finally, we have the public, who has an interest (perhaps more so than the other parties) because their lives are directly affected by the premiums charged.

It is interesting to note that, the modelers, who are most familiar with the models, tend to be same people designing and communicating the explanations. This poses a challenge that @millerExplainableAI2017 calls "inmates running the asylum", where the modelers design explanations for *themselves* rather than the intended audience.

@@

With the above considerations in mind, we propose a potential framework for interpreting ML models for insurance pricing. The actuarial profession, in collaboration with regulators, define a set of questions to be answered by explanations accompanying ML models, along with acceptance criteria and examples of successful explanations. In other words, interpretability for our purposes is defined as the ability of a model's explanations to answer the posed questions.

@@

While the realization of a set of such guidelines is an ambitious undertaking beyond the scope of this paper, we present in the next section a sample set of questions and techniques one can leverage to answer them. The goal of these case studies is twofold: to more concretely illustruate the proposed framework, and to expose the actuarial audience to modern ML interpretation techniques.

@@

(some definitions, e.g. global/model vs. local/instance, categorize questions accordingly)

## (answer each question)

(for each question, propose a technique, mention alternative techniques, pros/cons, implement, interpret results)

@@

(conclude)

>>>>

---
title: |
 Towards Explainability of Machine Learning Models in Insurance Pricing
type: WIP DRAFT 
author:
  - name: Kevin Kuo
    affil: a
    email: kevin@kasa.ai
  - name: Daniel Lupton
    affil: b
    email: dlupton@taylorandmulder.com
affiliation:
  - num: a
    address: |
      Kasa AI,
      3040 78th Ave SE #1271,
      Mercer Island, WA 98040, USA
  - num: b
    address: |
      Taylor & Mulder,
      10508 Rivers Bend Lane,
      Potomac, MD 20854, USA
bibliography: explain-pricing.bib
appendix: appendix.tex
citation_package: bibtex
abstract: |
  Abstract tbd
keywords: |
  actuarial science; general insurance
header-includes: |
  \usepackage{hyperref}
  \usepackage[utf8]{inputenc}
  \def\tightlist{}
output: 
  bookdown::pdf_book:
    base_format: rticles::tf_article
    number_sections: true
---

@@

## Not All Linear Models are Interpretable {#linear-models}

@@

Proposed by @hiltonConversationalProcesses1990 and interpreted by @millerExplanationArtificial2017 in the context of ML, model explanations are conversations or social interactions. One consequence of this identification is that explanations need to be *relevant* to the *audience*. In developing, filing, and operationalizing a pricing model, one needs to accommodate a variety of stakeholders, each of whom has a different set of questions, assumptions, and technical capacity. First, there are internal stakeholders at the company, which includes management and underwriters. While some of the individuals in this audience may be technical, they are likely less familiar with predictive modeling techniques than the actuaries and data scientists who build the models. Next, we have the regulators, who may have limited resources to review the models, and will focus on a specific list of questions motivated by statute. Finally, we have the public, who has an interest (perhaps more so than the other parties) because their lives are directly affected by the premiums charged.

It is interesting to note that, the modelers, who are most familiar with the models, tend to be same people designing and communicating the explanations. This poses a challenge that @millerExplainableAI2017 calls "inmates running the asylum", where the modelers design explanations for *themselves* rather than the intended audience. For example, they may be interested in technical questions like extrapolation behavior and shape the explanations accordingly, which may be lost on the prospective policyholder.

@@

With the above considerations in mind, we propose a potential framework for interpreting ML models for insurance pricing: the actuarial profession, in collaboration with regulators, define a set of questions to be answered by explanations accompanying ML models, along with acceptance criteria and examples of successful explanations. In other words, interpretability for our purposes is defined as the ability of a model's explanations to answer the posed questions.

@@

While the realization of a set of such guidelines is an ambitious undertaking beyond the scope of this paper, we present in the next section a sample set of questions and techniques one can leverage to answer them. The goal of these case studies is twofold: to more concretely illustrate the proposed framework, and to expose the actuarial audience to modern ML interpretation techniques.

@@

Now that we have established a framework for model interpretation in the form of asking and answering relevant questions, we demonstrate examples of such exchanges via an illustrative case study. Analytically, our starting point is a fitted deep neural network model for predicting loss costs. As the modeling details are of secondary importance, they are available in the appendix (TODO: add appendix and link to it.) The questions that we ask of the model are as follows:

1. What are the most important predictors in the model?
2. How does the predicted loss cost change, on average, as we change an input?
3. For a particular policyholder, how does each characteristic contribute to the loss cost prediction?

## A Simplified View of Interpretation Techniques

(TODO: add citation to interpretable ml book)

Before we dive into the answering questions, we provide some background on model interpretation techniques. As the literature is quite sizeable and fast moving, we do not attempt a full taxomony of techniques. The goal is to orient ourselves among broad categories of techniques, so we can map them to tasks indicated by the questions being asked. We classify model interpretation techniques across two dimensions: intrinsic vs. post-hoc and global vs. local.

### Intrinsic vs. Post-hoc

Intrinsic model interpretation draws conclusions from the structure of the fitted model and are what we typically associate with "interpretable" classes of models. This is only viable with models with simple structures, such as the sparse linear model and shallow decision tree we see in Section \ref{linear-models}, where we arrive at explanations by reading off parameter estimates or a few decision rules. For algorithms that produce models with complex structure that do not lend themselves easily to intrinsic exploration, we can appeal to post-hoc techniques. This class of techniques interrogate the model by presenting it with data for and observing the prediction behavior of the model. These techniques are concerned with only the inputs and outputs, and hence are *agnostic* of the model itself, which means they can be applied also to simple models. Since most useful ML models have a level of complexity beyond the threshold of intrinsic interpretability, we focus on model-agnostic techniques in our case study. As we will see later on, the data that we present to the models are usually some perturbed variations of test data.

### Global vs. Local

Along the other dimension, we categorize model interpretations as global, or model-level, and local, or instance-level. The former class provides insights with respect to the model as a whole. Some examples of these eplanations include variable importances and sensitivities, on average, of the predicted response with respect to individual predictors. In our case study, questions 1 and 2 are associated with this category. On the other hand, question 3 pertains to an individual prediction, which would fall in the local interpretation category. In addition to individual variable attribution, we can also inquire about what would happen to the current predicted response if we were to perturb specific predictor variables.

## Answering the Questions

Having aligned the questions with the categories of interpretation techniques, we now introduce a sample of appropriate techniques to answer them.

### Variable Importance

> "What are the most important predictors in the model?"

For linear models and their generalizations, and some ML models, measures of variable importance can be obtained from the fitted model structure. In the case of GLMs, one can observe the magnitudes of the $t$-statistics, whereas for random forests, one can use out-of-bag errors [@breimanRandomForests2001]. For more complex models, such as the neural network in our case study, we need to devise another approach.

We follow the methodology of permutation feature importance as described in @fisherAllModels2018, and utilize the notation introduced by @biecekPMVEE. The gist of the technique is as follows: to see how important a variable is, we make predictions without it and see how much worse off we are in terms of accuracy. One way to achieve this would be to re-fit the model many times (as many times as the number of variables.) However, this may be intractable with lengthy model training times or large number of variables, so a more popular approach is to instead keep the same fitted model but permute the values of each predictor.

More formally, let $y$ denote the vector of responses, $X$ denote the matrix of predictor variables, $\widehat{f}$ denote the fitted model, and $L = \mathcal{L}(\widehat{f}(X), y)$ denote the value of the loss function, which is mean squared error in the case of regression. Now, if $\widetilde{X}^{j}$ denotes the predictor matrix where the $j$th variable has been permutated, then we can compute the loss with the permuted dataset as $L^{-j} = \mathcal{L}(\widehat{f}(\widetilde{X}^{j}), y)$. With this, we define the absolute variable importance $VI^{j}$ as $L^{-j} - L$ or the relative variable importance as $VI_{R}^{j} = \frac{L^{-j}}{L}$.

In Figure \@ref(fig:fi), we show a plot of variable importances. We see that FOO and BAR are the most important variables in terms of contributing accuracy to our model. Note that these measures do not provide information regarding the directional sensitivity of the predictors on the response. In our particular example, we see that the "make" variable contributes most to the accuracy of the model with "XXX" contributing the least. This provides a way for the audience to quickly glance at the most relevant variables, and ask further question as necessary. For example, if a variable that can potentially be a proxy for a discriminatory variable appears as very important, more investigation would be warranted.

```{r fi, out.width='60%', fig.align = "center", fig.cap="Permutation feature importances for ML model.", echo = FALSE}
knitr::include_graphics("figures/fi-plot.png")
```

### Partial Dependence Plots

> "How does the predicted loss cost change, on average, as we change an input?"

For this question, we again consider first how it would be answered in the GLM setting. When the input predictor in question is continuous, we can answer the question by looking at the estimated coefficient, which provides the change in the response per unit change in the predictor (on the scale of the linear predictor). For non-parametric models and neural networks, where no coefficients are available, we can appeal to partial dependence plots (PDP), first proposed by @friedmanGreedyFunction2001 for gradient boosting machines (GBM).

To describe PDP, we first introduce some notation. Let $x^j$ denote the input variable of interest. Then we define the partial dependence function as

\begin{equation}
h(z) = E_{X^{-j}}[\widehat{f}(x|x^j = z)],
\end{equation}

where the expectation is taken over the distribution of the other predictor variables. In other words, we marginalize them out so we can focus on the relationship between the predicted response and the variable of interest. Empirically, we estimate $h$ by

\begin{equation}
\widehat{h}(z) = \frac{1}{N}\sum_{i = 1}^{N}\widehat{f}(x_i|x_i^{j} = z).
\end{equation}

In Figure \@ref(fig:pdp), we exhibit the PDP for the "vehicle age" variable. We see that the average predicted loss cost decreases with vehicle age until the latter is around 18. Note that the ticks on the axis tell us that the data is quite thin for vehicle age > 18, so the apparent upward trend to the right is driven by just a few data points. With this in mind, we can conclude that as cars get older, we expect the loss cost predictions generated by our model to decrease.

```{r pdp, out.width='60%', fig.align = "center", fig.cap="Partial dependence plot for ML model.", echo = FALSE}
knitr::include_graphics("figures/pdp-plot.png")
```

### Variable Attribution

> "For a particular policyholder, how does each characteristic contribute to the loss cost prediction?"

In the previous two examples, we look at model-level explanations; now we move on to one where we investigate one particular prediction instance. As before, we consider how we would approach the question for linear models. For a GLM with a log link common in ratemaking applications, for example, we simply start with the base rate, then the exponentiated coefficients would have multiplicative effects on the final rate. Similar to the previous examples, for ML models in general we do not have directly interpretable weights. Instead, one way to arrive at variable contributions is calculating the change in the expected model prediction for each predictor, conditioned on other predictors.

Formally, for a fitted model $\widehat{f}$ and a specific instance $x_*$, we would like to decompose the model prediction $\widehat{f}(x_*)$ into

\begin{equation}
\widehat{f}(x_* ) = v_0 + \sum_{j=1}^p v(j, x_*),
\end{equation}

where $v(j, x_*)$ denotes the contribution of the $j$th variable in instance $x_*$, defined as

\begin{equation}\label{eq:var-contrib}
v(j, x_* ) = E_X[\widehat{f}(X) | X^1 = x_* ^1, \dots, X^j = x_* ^j] - E_X[\widehat{f}(X^1 = x_* ^1, \dots, X^{j-1} = x_*^{j - 1})].
\end{equation}

Hence, the contribution of the $j$th variable to the prediction is the incremental change in the expected model prediction when we set $X^j = x_*^ j$ assuming the other variables take their values in $x_*$. Note here that this definition implies that the order in which we consider the variables affects the results. Empirically, the expectations in \@ref(eq:var-contrib) are calculated by sampling the test dataset.

In Figure \@ref(fig:breakdown) we exhibit a waterfall plot of variable contributions. The "intercept" value denotes the average model prediction. The predicted loss cost for this particular policyholder is slightly less than average; the characteristics that makes this policyholder more risky are the fact that he is a male between the ages of 18 and 25; counteracting the risky driver characteristics are the vehicle properties: it is a GM vehicle built domestically and is seven years old.

```{r breakdown, out.width='80%', fig.align = "center", fig.cap="Variable contribution plot for ML model.", echo = FALSE}
knitr::include_graphics("figures/breakdown-plot.png")
```

@@
